---
title: "activity_analysis"
author: "GabiK"
date: "5/20/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(glue)
library(scales)
library(ggpubr)
library(lme4)
library(DHARMa)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")

#N.B. after installing new packages, re-start R before running the script
```
# LIGHT BARRIER-BASED ACTIVITY
# Data preparation
```{r}
# activity files with weather data
frd=read.csv("data/Friedland_activity_LB.csv")
dem=read.csv("data/Demmin_activity_LB.csv")
pen=read.csv("data/Peenemuende_activity_LB.csv")
ank=read.csv("data/Anklam_activity_LB.csv")

merged=rbind(ank,dem,frd,pen)
merged$monthday=substr(merged$pseudodate, start=6, stop=10)
merged= merged %>% filter(!is.na(passes)) #remove days with NA passes
```
# Figure 1A in the manuscript
```{r}
# data for plotting
merged2=merged %>%
  group_by(site, Flash) %>%
  add_count(name="samplesize") %>%
  mutate(Flash=fct_recode(Flash,"flash-off" = "off", "flash-on"="on")) %>% 
  mutate(samplesize = glue("{ Flash } \n (N={ samplesize })"))
# activity on flash-on vs flash-off nights, faceted by site
flash_activity=ggplot(data = merged2, 
       aes(y = passes, x = samplesize , fill = Flash)) +
  geom_flat_violin(position = position_nudge(x = 0.25, y = 0), alpha = .8) +
  geom_point(aes(y = passes, color =Flash), 
             position = position_jitter(width = .15), size = .5, alpha = 0.8) +
  geom_boxplot(width = .3, outlier.shape = NA, alpha = 0.5) +
  expand_limits(x = 1) +
  facet_wrap(~site, scales="free")+
  coord_flip()+
  labs(x="",
       y="bat activity / night")+
  scale_color_manual(values=c("#4A6FE3", "#D33F6A"))+
  scale_fill_manual(values=c("#4A6FE3", "#D33F6A"))+
  scale_x_discrete(expand=c(0,0.25))+
  #scale_y_continuous(breaks = c(seq(0,300, by=50)))+
  theme_bw() +
  theme(legend.position = "none", axis.text.y = element_text(hjust=0.5))
flash_activity
```
# Figure 1B in the manuscript
```{r}
# activity and proportion of night hours with rain, faceted by site
rain_activity=merged %>% 
  filter(Flash!="NA") %>% 
  ggplot(aes(x=precProp,y=passes))+
  geom_point(aes(col=Flash))+
  scale_color_manual(values = c("#4A6FE3", "#D33F6A"))+
  geom_smooth(method="lm", data=merged[merged$Flash=="on",], col="#D33F6A", fill="#D33F6A", alpha=0.2)+
  geom_smooth(method="lm", data=merged[merged$Flash=="off",], col="#4A6FE3", fill="#4A6FE3", alpha=0.2)+
  labs(y="bat activity / night",
       x="proportion of hours with rain / night")+
  facet_wrap(~Site, scales="free")+
  theme_bw()+
  scale_x_continuous(labels = label_percent())+
  theme(legend.position = c(0.96, 0.9),
        legend.background = element_rect(fill = "#ffffffaa", colour = "black"),
        legend.key.size = unit(0.2,"line"),
        legend.title = element_text(size=10),
        legend.text = element_text(size=8))
rain_activity
```
# Figure 1C in the manuscript
```{r}
# activity and temperature at sunset, faceted by Site
temperature_activity=merged %>% 
  ggplot(aes(x=sunset_temp,y=passes))+
  geom_point(aes(col=Flash))+
  scale_color_manual(values = c("#4A6FE3", "#D33F6A"))+
  geom_smooth(method="lm", data=merged[merged$Flash=="on",], col="#D33F6A", fill="#D33F6A", alpha=0.2)+
  geom_smooth(method="lm", data=merged[merged$Flash=="off",],col="#4A6FE3", fill="#4A6FE3", alpha=0.2)+
  labs(y="bat activity / night",
       x="temperature at sunset (°C)")+
  facet_wrap(~Site, scales="free")+
  theme_bw()+
  theme(legend.position = c(0.96, 0.9),
        legend.background = element_rect(fill = "#ffffffaa", colour = "black"),
        legend.key.size = unit(0.2,"line"),
        legend.title = element_text(size=10),
        legend.text = element_text(size=8))
temperature_activity
```
# save plot
```{r}
plots=ggarrange(flash_activity, rain_activity,
          temperature_activity, nrow=3, ncol=1,
          labels=c("A","B","C"))
plots
# save high resolution TIFF file
ggplot2::ggsave(plots,file = "flash_weather_activity_raw.tiff",
                device='tiff', dpi=800,  width = 210, 
                height = 297,  units = "mm")
```
# Modelling light barrier-based bat activity
```{r}
#rescale continuous variables
merged$precProp_stand=scale(merged$precProp, center = F)
merged$sunset_temp_stand=scale(merged$sunset_temp, center = F)

#calculate correlation coefficients
merged %>% dplyr::select(Week, precProp_stand, sunset_temp_stand) %>% cor()

merged$Week=as.factor(merged$Week)

#final model
model=glmer.nb(passes ~ precProp_stand + sunset_temp_stand + Flash + (1|Site),data=merged) 
#model output              
summary(model)
```

```{r}
#calculate odds ratio with confidence interval
cc=confint(model,parm="beta_")  
ctab=cbind(est=fixef(model),cc)
ctab=as.data.frame(ctab)
ctab=setNames(cbind(rownames(ctab), ctab, row.names = NULL), 
              c("variable","odds ratio","lcl","ucl"))
ctab$variable=c("intercept", "proportion of \n rainy night hours \n (standardized)","air temperature \n at sunset \n (standardized)", "flash on")
ctab[,2:4] <- exp(ctab[,2:4])
ctab
```
# Model validation
```{r}
set.seed(42)
testUniformity(model) #tests if the overall distribution conforms to expectations
testDispersion(model) # tests if the simulated dispersion is equal to the observed dispersion
testQuantiles(model) #fits a quantile regression
```

# VIDEO-BASED ACTIVITY
# Data preparation
```{r}
# activity files with weather data
frd=read.csv("data/Friedland_activity_video.csv")
dem=read.csv("data/Demmin_activity_video.csv")
pen=read.csv("data/Peenemuende_activity_video.csv")
ank=read.csv("data/Anklam_activity_video.csv")

merged=rbind(ank,dem,frd,pen)
merged$monthday=substr(merged$pseudodate, start=6, stop=10)
merged= merged %>% filter(!is.na(passes)) #remove days with NA passes
```
# Figure S3A in the manuscript
```{r}
# data for plotting
merged2=merged %>%
  group_by(site, Flash) %>%
  add_count(name="samplesize") %>%
  mutate(Flash=fct_recode(Flash,"flash-off" = "off", "flash-on"="on")) %>% 
  mutate(samplesize = glue("{ Flash } \n (N={ samplesize })"))
# activity on flash-on vs flash-off nights, faceted by site
flash_activity=ggplot(data = merged2, 
       aes(y = passes, x = samplesize , fill = Flash)) +
  geom_flat_violin(position = position_nudge(x = 0.25, y = 0), alpha = .8) +
  geom_point(aes(y = passes, color =Flash), 
             position = position_jitter(width = .15), size = .5, alpha = 0.8) +
  geom_boxplot(width = .3, outlier.shape = NA, alpha = 0.5) +
  expand_limits(x = 1) +
  facet_wrap(~site, scales="free")+
  coord_flip()+
  labs(x="",
       y="bat activity / night")+
  scale_color_manual(values=c("#4A6FE3", "#D33F6A"))+
  scale_fill_manual(values=c("#4A6FE3", "#D33F6A"))+
  scale_x_discrete(expand=c(0,0.25))+
  theme_bw() +
  theme(legend.position = "none", axis.text.y = element_text(hjust=0.5))
flash_activity
```
# Figure S3B in the manuscript
```{r}
# activity and proportion of night hours with rain, faceted by site
rain_activity=merged %>% 
  filter(Flash!="NA") %>% 
  ggplot(aes(x=precProp,y=passes))+
  geom_point(aes(col=Flash))+
  scale_color_manual(values = c("#4A6FE3", "#D33F6A"))+
  geom_smooth(method="lm", data=merged[merged$Flash=="on",], col="#D33F6A", fill="#D33F6A", alpha=0.2)+
  geom_smooth(method="lm", data=merged[merged$Flash=="off",], col="#4A6FE3", fill="#4A6FE3", alpha=0.2)+
  labs(y="bat activity / night",
       x="proportion of hours with rain/ night")+
  facet_wrap(~Site, scales="free")+
  theme_bw()+
  scale_x_continuous(labels = label_percent())+
  theme(legend.position = c(0.96, 0.9),
        legend.background = element_rect(fill = "#ffffffaa", colour = "black"),
        legend.key.size = unit(0.2,"line"),
        legend.title = element_text(size=10),
        legend.text = element_text(size=8))
rain_activity
```
# Figure S3C in the manuscript
```{r}
# activity and temperature at sunset, faceted by Site
temperature_activity=merged %>% 
  ggplot(aes(x=sunset_temp,y=passes))+
  geom_point(aes(col=Flash))+
  scale_color_manual(values = c("#4A6FE3", "#D33F6A"))+
  geom_smooth(method="lm", data=merged[merged$Flash=="on",], col="#D33F6A", fill="#D33F6A", alpha=0.2)+
  geom_smooth(method="lm", data=merged[merged$Flash=="off",],col="#4A6FE3", fill="#4A6FE3", alpha=0.2)+
  labs(y="bat activity / night",
       x="temperature at sunset (°C)")+
  facet_wrap(~Site, scales="free")+
  theme_bw()+
  theme(legend.position = c(0.96, 0.9),
        legend.background = element_rect(fill = "#ffffffaa", colour = "black"),
        legend.key.size = unit(0.2,"line"),
        legend.title = element_text(size=10),
        legend.text = element_text(size=8))
temperature_activity
```
# save plot
```{r}
plots=ggarrange(flash_activity, rain_activity,
          temperature_activity, nrow=3, ncol=1,
          labels=c("A","B","C"))
plots
# save high resolution TIFF file
ggplot2::ggsave(plots,file = "flash_weather_activity_corrected.tiff",
                device='tiff', dpi=800,  width = 210, 
                height = 297,  units = "mm")
```
# Modelling video-based bat activity
```{r}
#rescale continuous variables
merged$precProp_stand=scale(merged$precProp, center = F)
merged$sunset_temp_stand=scale(merged$sunset_temp, center = F)

#calculate correlation coefficients
merged %>% dplyr::select(Week, precProp_stand, sunset_temp_stand) %>% cor()

merged$Week=as.factor(merged$Week)

#final model
model=glmer.nb(passes ~ precProp_stand + sunset_temp_stand + Flash +  
                      (1|Site),data=merged) 
#model output              
summary(model)
```

```{r}
#calculate odds ratio with confidence interval
cc=confint(model,parm="beta_")  
ctab=cbind(est=fixef(model),cc)
ctab=as.data.frame(ctab)
ctab=setNames(cbind(rownames(ctab), ctab, row.names = NULL), 
              c("variable","est","lcl","ucl"))
ctab$variable=c("intercept", "proportion of \n rainy night hours \n (standardized)","air temperature \n at sunset \n (standardized)", "flash on")
ctab[,2:4] <- exp(ctab[,2:4])
ctab
```
# Model validation
```{r}
set.seed(42)
testUniformity(model) #tests if the overall distribution conforms to expectations
testDispersion(model) # tests if the simulated dispersion is equal to the observed dispersion
testQuantiles(model) #fits a quantile regression
```

